---
title: "Build a k-means pallette from a png image"
author: "John Minter"
date: "Last updated: 2018-06-15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Build a k-means pallette from a png image}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(rAnaLab))))
```

This function was developed from an example from
[www.milanor.net](http://www.milanor.net/blog/build-color-palette-from-image-with-paletter/).
We load a **png** image of a color pallette called `rocks-in-the-river` from
(www.color-hex.com)[http://www.color-hex.com/color-palette/61776].
We first load a pallette image from our system data and then call the
function. The function returns an image with swatches with
the hex values and a list of the hex values.

```{r, comment=NA, fig.width=4, fig.align='center'}
library(rAnaLab)
fi <- system.file("extdata", "rocks-in-the-river.png", package = "rAnaLab")
res <- kmeans_pallette(fi)
res
```

The values from the site are:

```{r}
rocks_int_the_river <- c('#4e5166','#7c90a0','#b5aa9d','#b9b7a7','#747274')
```

So we obtained the expected answer.
