---
title: "Lognormal Distribution Plot"
author: "John Minter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lognormal Distribution Plot}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

First load the test data

```{r loadData, message=FALSE}
library(rAnaLab)
library(pander)
data(diam)
pander(summary(diam))
```

We have `r nrow(diam)` individual measurements of the equivalent
circular diameter of a silver halide photographic emulsion. To be
consistent with the usual colloid science terminology, we will
henceforth use the term **dispersion**.  These were collected using a
Gatan slow scan CCD camera in a FEI CM20T TEM using a Gatan cooling
holder at -170C. This minimizes radiolysis of the AgX particles.

Next we bin the data and look at the output. We want to use a reference
bin diameter of 10.0 nm and specify the 8^th^ root of 2..

```{r}
l.b <- rb.lognormal.bin(diam[,1], n.root.2=8.0,  base.val=10.0)

print(str(l.b))
```
The function outputs a list that contains three elements:

1. The first item is a R dataframe. The dataframe contains four vectors:

- The log of the log of the diameter (logx). In this case it is the 
**midpoint** of the bin.
- The diameter (x). Again, the midpoint of the bin.

- The counts (cts), the number of features measured that were compliant
with the agglomeration criteria.

- The density (measured from R's histogram). In this case it is the
fraction of the particles in the bin.

2. The second item in the list is the mean log diameter. For a single
mode log-normal distribution, the **geometric mean diameter** is 
`exp(mean(log(d)))`.

3. The third item in the list is the standard deviation of the log
diameter. For a single mode log-normal distribution, the
**geometric standard diameter** is `exp(sd(log(d)))`.


next, we want to create a data frame for plotting with `ggplot2`.


```{r, makeDataFrame}
binned <- make.log.bin.df(l.b)
pander(head(binned))
```

And look at the high end

```{r, highDiam}
pander(tail(binned))
```

We want to get a first view of a plot. There are problems passing
`waver()` to `ggplot2` from a package, so I developed the function
`gp.logx.distn.tst` to use mostly defaults.

```{r, testPlot, fig.width=7}
plt <- ggp.logx.distn.tst(binned,
                          "AgX Grain Diameter",
                           x.units=-9)

print(plt)
```

Finally, we can pass all the controls to the `ggp.logx.distn` for a 
final plot


```{r, finalPlot, fig.width=7}
plt <- ggp.logx.distn(binned,
                      "AgX Grain Diameter",
                       x.units=-9,
                       d.max=0.025,
                       limits=c(50, 500),
                       breaks=c(50, 100, 500),
                       minor_breaks=c( 60, 70, 80, 90, 200, 300, 400))

print(plt)
```
