---
title: "Simulating Lognormal Particle Size Distributions"
author: "John Minter"
date: "Last updated: 2017-10-25"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating Lognormal Particle Size Distributions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(rAnaLab))))
```
We want to simulate a lognormal particle size distribution and
demonstrate the properties of the distribution. First, load the
libraries we need and demonstrate the generation of the breaks for
lognormal binning of a real dataset. We will plot the separation of
the bin breaks.

```{r loadLibraries, fig.width=7, fig.height=3, fig.cap="Note the increased separation between the bin breaks"}
library(rAnaLab)
library(pander)

data(diam)
x <- calc.ln.bin.breaks(diam[,1], n.root.2=4, base.val=0.1)
y <- rep(1,length(x))
df <- data.frame(x=x, y=y)
plt <- ggplot() +
           geom_point(data=df, aes(x=x, y=y),
                      colour="darkblue") +
           ylim(0.75, 1.25) +
           xlab("bin break position [nm]") +
           ylab("") 
print(plt)
```

Let's simulate a single mode lognormal distribution. We start by setting
variables for the **geometric mean diameter** (`gmd`) and the
**geometric standard deviation** (`gsd`). We will bin the results
with bins separated by the **8^th^ root of 2** and with a
**reference bin break** at 0.1 nm. First we will take a peek at the
structure of the output that `sim.lognormal()` produces.

```{r simulateLognormal}
num <- 100000
gmd <- 50
gsd <- 2.0

sim <- sim.lognormal(gmd, gsd, num, 8, 0.1)

print(str(sim))
```

Note that we have a dataframe with entries for the diameter (in this
case the **midpoint** of the bin), the associated counts (the
**number of particles in each bin**), and the density (the 
**fraction of particles** in each bin). This is prety much everything
that we need for a plot and subsequent analysis.

Let's calulate some undelying parameters of the lognormal distribution
and plot the results.

```{r, plotDistribution, fig.width=7, fig.cap="A single mode lognormal distribution. The **mode** is shown in red, the **geometric mean diameter** is shown in blue."}
mu <- log(gmd)
sigma <- log(gsd)
mode <- exp(mu - sigma^2)
lt <- 1.1

diaPlt <- ggplot() +
              geom_point(data=sim, aes(x=diam, y=dens), colour="darkblue") +
              scale_x_log10() +
              xlab("Equivalent Circular Diameter (ECD) [nm]") +
              ylab("density") +
              ggtitle(sprintf("%s (n=%d)", "lognormal distribution", num)) +
              theme(axis.text=element_text(size=12),
                   axis.title=element_text(size=12),
                   plot.title=element_text(hjust = 0.5)) +
              geom_vline(xintercept=gmd, color="blue", size=lt) + 
              geom_vline(xintercept=mode, color="red", size=lt)

print(diaPlt)

```

Note that the **mode** of the distribution (the peak) is **below**
the value of the **geometric mean diameter**, i.e. the point above
which **half the particles occur**.

Finally, let's print out a table

```{r makeTable, results='asis'}
iDigits <- 4
feature <- c("mu",
             "sigma",
             "mode", 
             "gmd",
             "gsd")
value <- c(sprintf("%.4f", round(mu,iDigits)),
           sprintf("%.4f", round(sigma, iDigits)),
           sprintf("%.4f", round(mode, iDigits)),
           sprintf("%.4f", round(gmd, iDigits)),
           sprintf("%.4f", round(gsd, iDigits)))

res <- data.frame(feature=feature, value=value)
pandoc.table(res, justify = c('left', 'right'))
```


